Relict deciduous oaks show high resilience to drought despite the effect of land-use legacies

### 2019-07-13 man/ecosystems/ms_ecosystems.Rmd
update ms

## **Author Contributions**
AJPL conceived and desing the study. AJPL and GGI conducted lab work and led article preparation. 
AJPL, GGI and RZ conducted field work, collect the data and contribute to writing and editing the manuscript. All authors contribute to the drafts and gave final approval for publication.

@@

In addition, the forest and managment history of our sampling sites was inferred from a detailed analysis of historical land-use changes. For this purpose, an exhaustive review of historical documents was done to compile information about socio-economical activities affecting forest Table XX

$ convertir en tabla $ 
: land-uses, forest managment actions, mining, wildfires, etc. Several documentary sources were reviewed: historical documents and maps [*e.g.* @Titos1990]; detailed mining reports [*e.g.* @Maestre1858]; official information about recent wildfires events and forest management practices [@Bonet2016obsnev_forest]; livestock farming [*e.g.* @MorenoLlorca2016]; traditional irrigation ditches [@RuizRuiz2017] and other studies reviewing the socioeconomic dynamics of forest of Sierra Nevada at different scales [@JimenezOlivencia2015; @MorenoLlorca2016].

@@

Greenness and tree-growth of *Q. pyrenaica* showed different sensitivity to severe droughts. During the last drought events EVI was reduced to 90.25 % (Table 1), whereas BAI was reduced to 70.83 % (Table S2), suggesting a lower sensitivity of the greenness than tree-growth to drought, particularly for sites under drier climate. For our driest site (SJ, the northern one), BAI were reduced to 44.5 % respect to that of the preceding period during 2005 drought event, whereas the EVI was reduced to 81.9 % (Tables S1 and S2). These findings are in accordance with previously works which showed that tree-growth are more sensitive metrics of forest resilience than net primary productivty [*e.g.* @Babst2013; @Coulthard2017; @Gazol2018]. Besides the different scales of both approaches, *i.e.* tree-based *versus* pixel-based; the growth reduction seems to be more mediated by sink that by source limitations [@Gazol2018]. Response of greenness to drought not only reflects response of tree, because remote sensing indices (EVI) captures signals from all vegetation covered by the pixel (not only tree but also understory vegetation) [see @Dionisioetal2012IPoGEC]. On the other hand, tree-ring width data provide an accurate measure of growth responses to droughts [@Bhuyan2017; @Gazol2018].

@@

The Mediterranean mountains, even at high elevations, have suffered strongly transformation of the landscapes driven by human activities throughout history. Multiple evidence from palaeoecological studies indicated an intense human impact on vegetation of Sierra Nevada since 3000 cal year BP [@GarciaAlix2017]. From that moment onwards, increases on fire occurrence and both grazing and mining activities, were recorded for this mountain region, with an sharp intensification of the human activities in the last 150 years [@GarciaAlix2017]. Mining, forest clearings, fuelwood and charcoal exploitations, pastoralism and also wars, have strongly impacted on the forest resources in Sierra Nevada (Table S4). As a consequence of all those activities a loss of about 90% of broadleaf *Quercus* species cover from medium and low elevation occurred in this mountain region [@JimenezOlivencia2015].

@@

We found similar tree competence levels in our sampled sites (Table 1), but differences in tree size and age suggest different management origin (*i.e.* land-use legacy). The review of historical documents revealed that our study sites had different land-use trajectories driven mainly by the differential pattern of the natural resource use [@JimenezOlivencia2015; @MartinCivantos2014]. On the one hand, the northern sites (*e.g.* San Juan), show a sequentially distribution of the land uses along the elevational gradient [@JimenezOlivencia2015]: grasslands and shrublands for cattle farming are located at high elevations; then forests formation with some croplands; and at lowlands, irrigated terraces with tree crops. Other activities (*e.g.* mining) could also determine the use of natural resources, and therefore the forest structure. For instance, the SJ oak woodland is located in an area with a high concentration of mines and quarries that have been exploited intermitently throughout history. Historical documents indicated two periods of intense mining activity: the second half of the 19th century after the publication of detailed mineralogical reports by [@Maestre1858]; and during the first decades of the twentieth century until 1960, which is the last year in which there is evidence of the existence of mining activity. On the other hand, woodland areas of the southern slopes (Cáñar) are mixed with a greater percentage of croplands, even reached high elevation (mainly barley, rye and potatoes) [@Calatrava2019]. Forest resources, like firewood, charcoal and acorns, have been continuously exploited in southern sites through history [@Valbuena2013; @JimenezSerrano2004]. In Cáñar site, there were an arraigated tradition of charcoal extraction since fifteenth century [@JimenezSerrano2004] up to middle of the 20th century. From this moment to the present, there were a sharp decrease of the wood extraction, mainly due to rural abandonment [@MesaTorres2009; @Bonet2014_conama]. Acorn resources have been tradiotionally exploited in this location up to 1950s [@MesaTorres2009; @Bonet2014_conama].

Those different patterns of resource uses can help to understand the tree-growth changes recorded in our chronologies (Figure 5). We observed a tree-growth release event at 1940s in the SJ oak woodland (Figures 5, S4) which concurs with one period of maximum mining activity for this area. Several documentary sources indicate an apogee of the mining activity during 1925 to 1957 period, supported by both the improvement of the paths to transport the mineral and by the creation of new railway connections [@Titos1990]. During this period there were an increase on the use of timbers for the tunnels of the mines and several furnaces that required great quantities of fuelwood to melt the mineral were active in this area [@Titos1990]. This heavily exploitation of the forest resources could affect to a major part of this oak woodland, since percentage of trees affected by GC > 50 % reaches values above 50 % (Figure S4). Our results also coincide with a concurrent but less intense tree-growth release event reported for a closed oak woodland [@GeaIzquierdo2014]. In addition, paleoecological studies carried out in nearby alpine bogs have recorded increases in the heavy metals concentration since the end of 18th century until mid-20th, which coincides with the maximum activity of the mining in this area [@GarciaAlix2017].

The other release event observed for SJ site during 1995-2000 was lower than the occurred at 1940, but affecting more trees (Figures 5, S4). We revised the forest practices carried out in this area in the last 30 years [@Bonet2016obsnev_forest], but we did not find any clearing or cutting event during 1995 - 2000 period. Two non-mutually exclusive ways could explain the tree-growth release observed. Firstly it could be related with a natural drought-induced mortality event after 1995, as was reported for Mediterranean tree species [*e.g.* @Penuelas2000; @Lloret2004]. On the other hand, we obtained strong positive correlations of SPEI with tree-growth for this site (Figure 8), which suggests a high sensitivity to water availability of this slightly more xeric site (Table 1) [@GeaIzquierdo2014]. Previous works reported a sharp decline (or non-production) in latewood production during extreme drought events [@Corcuera2006]. Since latewood are less vulnerable to embolism than earlywood vessels [@Corcuera2006], the reduction or even the non-production of former could negatively affect to tree-growth and also enhanced the mortality, particularly for drier sites [@Corcuera2006; @GeaIzquierdo2014].

The only positive release events showed for CA-High site could be related with conversion from closed forest to an open silvopastoral system, a common management applied in the past [@Canellas2004; @GeaIzquierdo2014] documented in this site [@Valbuena2013].

>>>>

## **Author Contributions**
AJPL conceived and desing the study and wrote the manuscript. AJPL and GGI conducted lab work and led article preparation. 
AJPL, GGI and RZ conducted field work, collect the data and contribute to writing and editing the manuscript. All authors contribute to the drafts and gave final approval for publication.

@@

In addition, the forest and managment history of our sampling sites was inferred from a detailed analysis of historical land-use changes. For this purpose, an exhaustive review of historical documents was done to compile information about socio-economical activities affecting forest (Table S4).

```{r eval=FALSE, echo=FALSE}
# : land-uses, forest managment actions, mining, wildfires, etc. Several documentary sources were reviewed: historical documents and maps [*e.g.* @Titos1990]; detailed mining reports [*e.g.* @Maestre1858]; official information about recent wildfires events and forest management practices [@Bonet2016obsnev_forest]; livestock farming [*e.g.* @MorenoLlorca2016]; traditional irrigation ditches [@RuizRuiz2017] and other studies reviewing the socioeconomic dynamics of forest of Sierra Nevada at different scales [@JimenezOlivencia2015; @MorenoLlorca2016].
```

@@

Greenness and tree-growth of *Q. pyrenaica* showed different sensitivity to severe droughts. During the last drought events EVI was reduced to 90.25 % (Table 1), whereas BAI was reduced to 70.83 % (Table S2), suggesting a lower sensitivity of the greenness than tree-growth to drought, particularly for sites under drier climate. For our driest site (SJ, the northern one), BAI were reduced to 44.5 % respect to that of the preceding period during 2005 drought event, whereas the EVI was reduced to 81.9 % (Tables S1 and S2). These findings are in accordance with previously works which showed that tree-growth are more sensitive metrics of forest resilience than net primary productivty [*e.g.* @Babst2013; @Coulthard2017; @Gazol2018]. Besides the different scales of both approaches, *i.e.* tree-based *versus* pixel-based; the growth reduction seems to be more mediated by sink that by source limitations [@Gazol2018]. Response of greenness to drought not only reflects response of tree, because remote sensing indices (EVI) captures signals from all vegetation covered by the pixel (not only tree but also understory vegetation) [see @Dionisioetal2012IPoGEC]. In this sense, recent results indicate that remote sensing models underestimate the impact of drought on primary production [@Stocker2019], as soil moisture is not taken into account by this satellite estimates. This is particularly important for *Q. pyrenaica* forests, as soil moisture is a key growth-related factor for this species [@GeaIzquierdo2014]. On the other hand, tree-ring records allow to contextualize remote sensing data by reflecying climate or disturbance induced tree-growth anomalies over decades to centuries [@Babst2017] and provide an accurate measure of growth responses to droughts [@Bhuyan2017; @Gazol2018]. Combined anlaysis, such the approach used in our study, may also allow climate-induced variability in forest growth to be disentangled from that driven by community-level ecological processes [@Babst2017], since spatio temporal patterns of tree-ring measurement and earth observations (*i.e.* EVI) reflects processes at different spatio-temporal scales [Babst2018].

@@

The Mediterranean mountains, even at high elevations, have suffered strongly transformation of the landscapes driven by human activities throughout history. Multiple evidence from palaeoecological studies indicated an intense human impact on vegetation of Sierra Nevada since 3000 cal year BP [@GarciaAlix2017]. From that moment onwards, increases on fire occurrence and both grazing and mining activities, were recorded for this mountain region, with an sharp intensification of the human activities in the last 150 years [@GarciaAlix2017]. Mining, forest clearings, fuelwood and charcoal exploitations, pastoralism and also wars, have strongly impacted on the forest resources in Sierra Nevada, and on Pyrenean oak forest, as can be derived from the results of the review of historical documents (Table S4). As a consequence of all those activities a loss of about 90% of broadleaf *Quercus* species cover from medium and low elevation occurred in this mountain region [@JimenezOlivencia2015].

@@

We found similar tree competence levels in our sampled sites (Table 1), but differences in tree size and age suggest different management origin (*i.e.* land-use legacy). The analysis of growth changes revealed differences in forest history between sites (Figures 5 and S4), driven mainly by the differential pattern of the natural resource use [@JimenezOlivencia2015] (Table S4). On the one hand, the northern sites (*e.g.* San Juan), show a sequentially distribution of the land uses along the elevational gradient [@JimenezOlivencia2015]: grasslands and shrublands for cattle farming are located at high elevations; then forests formation with some croplands; and at lowlands, irrigated terraces with tree crops. Other activities (*e.g.* mining) could also determine the use of natural resources, and therefore the forest structure. For instance, the SJ oak woodland is located in an area with a high concentration of mines and quarries that have been exploited intermitently throughout history (Table S4). On the other hand, woodland areas of the southern slopes (Cáñar) are mixed with a greater percentage of croplands, even reached high elevation [@Calatrava2019]. Forest resources, like firewood, charcoal and acorns, have been continuously exploited in southern sites through history [@Valbuena2013]. In Cáñar site, there were an arraigated tradition of charcoal extraction since fifteenth century [@JimenezSerrano2004] up to middle of the 20th century. From this moment to the present, there were a sharp decrease of the wood extraction, mainly due to rural abandonment. Acorn resources have been tradiotionally exploited in this location up to 1950s (Tabla S4).

Those different patterns of resource uses can help to understand the tree-growth changes recorded in our chronologies (Figure 5). We observed a tree-growth release event at 1940s in the SJ oak woodland (Figures 5, S4) which concurs with one period of maximum mining activity for this area (1925 to 1957) (Table S4). During this period there were an increase on the use of timbers for the tunnels of the mines and several furnaces that required great quantities of fuelwood to melt the mineral were active in this area [@Titos1990]. This heavily exploitation of the forest resources could affect to a major part of this oak woodland, since percentage of trees affected by GC > 50 % reaches values above 50 % (Figure S4). Our results also coincide with a concurrent but less intense tree-growth release event reported for a closed oak woodland [@GeaIzquierdo2014]. In addition, paleoecological studies carried out in nearby alpine bogs have recorded increases in the heavy metals concentration since the end of 18th century until mid-20th, which coincides with the maximum activity of the mining in this area [@GarciaAlix2017]. The other release event observed for SJ site during 1995-2000 was lower than the occurred at 1940, but affecting more trees (Figures 5, S4). We revised the forest practices carried out in this area in the last 30 years [@Bonet2016obsnev_forest], but we did not find any clearing or cutting event during 1995 - 2000 period. Two non-mutually exclusive ways could explain the tree-growth release observed. Firstly it could be related with a natural drought-induced mortality event after 1995 (*personal communications*) as was reported for Mediterranean tree species [*e.g.* @Penuelas2000; @Lloret2004]. On the other hand, we obtained strong positive correlations of SPEI with tree-growth for this site (Figure 8), which suggests a high sensitivity to water availability of this slightly more xeric site (Table 1) [@GeaIzquierdo2014]. Previous works reported a sharp decline (or non-production) in latewood production during extreme drought events [@Corcuera2006]. Since latewood are less vulnerable to embolism than earlywood vessels [@Corcuera2006], the reduction or even the non-production of former could negatively affect to tree-growth and also enhanced the mortality, particularly for drier sites [@Corcuera2006; @GeaIzquierdo2014].

For southern sites, the only positive release events showed for CA-High site could be related with conversion from closed forest to an open silvopastoral system, a common management applied in the past [@Canellas2004; @GeaIzquierdo2014] documented in this site [@Valbuena2013].

### 2019-07-13 man/figures/resilience_long_term.Rmd
update ms

```{r plot_rt}
sevRt <- plotrel(sdf, v = "rt", x = 'd_severity', label = 'disturb_year',
                 coordX_eq = 6, coordY_eq = 0.5, titulo = "Rt") +
  xlab('') + ylab('') +
  theme(plot.margin = margin(r=-0.7, l=0, t=0, b=0, "pt"))

```{r plot_rs}
sevRs <- plotrel(sdf, v = "rs", x = 'd_severity', label = 'disturb_year',
                 coordX_eq = 6, coordY_eq = 0.5, titulo = "Rs") + 
  xlab('') + ylab('') +
  theme(plot.margin = margin(r=0, l=-0.7, t=0, b=0, "pt"))
```

```{r plot_rc}
sevRc <- plotrel(sdf, v = "rc", x = 'd_severity', label = 'disturb_year',
                 coordX_eq = 6, coordY_eq = 0.5, titulo = "Rc") +
  xlab("Drought Severity") + ylab('') +
  theme(plot.margin = margin(r=-0.7, l=-0.7, t=0, b=0, "pt"))
```

```{r plot_join, eval=FALSE, echo=FALSE}
res_long_term <- sevRt + sevRc + sevRs + 
  plot_layout(nrow = 1) 
```

@@

  ggplot(df, 
         aes_string(y=paste0(v, "_mean_mean"), x=x, label=label, colour=site, fill=site)) +
    geom_smooth(method='lm', se=FALSE, aes(group=1), colour='gray',  size = 1) + # linetype = "dotted",
    geom_smooth(method='lm', se=FALSE) + 
    geom_point(size=1.5, pch = 21) + 
    theme_bw() + 
    geom_text_repel(nudge_x = 0.5, size = 2, show.legend = FALSE) + 
    ggtitle(titulo) + 
    theme(plot.title = element_text(size = 11, face = "bold.italic"),
          panel.grid = element_blank(), 
          axis.text.y = element_text(size=9),
          axis.text.x = element_text(size=9),
          axis.title = element_text(size = 10)) +
    geom_text_repel(data = eqns, aes(x = coord_eqX, y = Inf, label = equation),
                    direction = "y", segment.size = 0, hjust = 0,
                    parse = TRUE, show.legend = FALSE) +
    annotate("text", x = .0, y = 10, 
             label=lm_eqn(lm(formulita, df)), parse=TRUE, 
             size = 5, fontface = "bold", hjust="left") 
}

@@

sevRt_site <- plotrelsite(sdfsite, v = "rt", x = 'd_severity', label = 'disturb_year', site = "site",
                 titulo = "Rt", coord_eqX = Inf) + xlab('') + ylab('')

+ 
  xlim(5,25)

sevRs_site <- plotrelsite(sdfsite, v = "rs", x = 'd_severity', label = 'disturb_year', site = "site",
                 titulo = "Rs", coord_eqX = -Inf) + xlab('') + ylab('') + xlim(5,25)

sevRc_site <- plotrelsite(sdfsite, v = "rc", x = 'd_severity', label = 'disturb_year', site = "site",
                 titulo = "Rc",  coord_eqX = -Inf) +
  xlab("Drought Severity") + ylab("") + xlim(5,25)

@@

ggsave(filename = here::here("./man/figures/svg/fig_6.jpg"), 
       dpi = 300,  plot=res_long_term_sites, width=8.5, height=7)

@@

sevRt_site95 <- plotrelsite(sdfsite95, v = "rt", x = 'd_severity', label = 'disturb_year', site = "site",
                 titulo = "Rt", coord_eqX = Inf) + xlab('') + ylab('')

sevRs_site95 <- plotrelsite(sdfsite95, v = "rs", x = 'd_severity', label = 'disturb_year', site = "site",
                 titulo = "Rs", coord_eqX = -Inf) + xlab('') + ylab('')

sevRc_site95 <- plotrelsite(sdfsite95, v = "rc", x = 'd_severity', label = 'disturb_year', site = "site",
                 titulo = "Rc",  coord_eqX = -Inf) + xlab("Drought Severity") + ylab("")

@@

ggsave(filename = here::here("./man/figures/svg/fig_6_95.jpg"), 
       dpi = 300,  plot=res_long_term_sites95, width=8.7, height=7)

>>>>

```{r, eval=FALSE, echo=FALSE}}
sevRt <- plotrel(sdf, v = "rt", x = 'd_severity', label = 'disturb_year',
                 coordX_eq = 6, coordY_eq = 0.5, titulo = "Rt") +
  xlab('') + ylab('') +
  theme(plot.margin = margin(r=-0.7, l=0, t=0, b=0, "pt"))

sevRs <- plotrel(sdf, v = "rs", x = 'd_severity', label = 'disturb_year',
                 coordX_eq = 6, coordY_eq = 0.5, titulo = "Rs") + 
  xlab('') + ylab('') +
  theme(plot.margin = margin(r=0, l=-0.7, t=0, b=0, "pt"))

sevRc <- plotrel(sdf, v = "rc", x = 'd_severity', label = 'disturb_year',
                 coordX_eq = 6, coordY_eq = 0.5, titulo = "Rc") +
  xlab("Drought Severity") + ylab('') +
  theme(plot.margin = margin(r=-0.7, l=-0.7, t=0, b=0, "pt"))

res_long_term <- sevRt + sevRc + sevRs + 
  plot_layout(nrow = 1) 
```

@@

  ggplot(df, 
         aes_string(y=paste0(v, "_mean_mean"), x=x, label=label, colour=site, fill=site)) +
    geom_smooth(method='lm', se=FALSE, aes(group=1), colour='gray',  size = 1) + # linetype = "dotted",
    geom_smooth(method='lm', se=FALSE) + 
    geom_point(size=1.5, pch = 21) + 
    theme_bw() + 
    geom_text_repel(nudge_x = 0.5, size = 2, show.legend = FALSE) + 
    ggtitle(titulo) + 
    theme(plot.title = element_text(size = 11, face = "bold.italic"),
          panel.grid = element_blank(), 
          axis.text.y = element_text(size=9),
          axis.text.x = element_text(size=9),
          axis.title = element_text(size = 10)) +
    geom_text_repel(data = eqns, aes(x = coord_eqX, y = Inf, label = equation),
                    direction = "y", segment.size = 0, hjust = 0,
                    parse = TRUE, show.legend = FALSE) +
    annotate("text", x = xLM, y = yLM, 
             label=lm_eqn(lm(formulita, df)), parse=TRUE, 
             size = 4, fontface = "bold", hjust="left", col = "gray40") 
}

@@

sevRt_site <- plotrelsite(sdfsite, v = "rt", x = 'd_severity', label = 'disturb_year', site = "site",
             titulo = "Rt", coord_eqX = Inf, xLM = -Inf, yLM = 0.5) + xlab('') + ylab('') + xlim(5,25)

sevRs_site <- plotrelsite(sdfsite, v = "rs", x = 'd_severity', label = 'disturb_year', site = "site",
                 titulo = "Rs", coord_eqX = -Inf, xLM = 10, yLM = 0.5) + xlab('') + ylab('') + xlim(5,25)

sevRc_site <- plotrelsite(sdfsite, v = "rc", x = 'd_severity', label = 'disturb_year', site = "site",
                 titulo = "Rc",  coord_eqX = -Inf, xLM = 10, yLM = 0.5) +
  xlab("Drought Severity") + ylab("") + xlim(5,25)

@@

ggsave(filename = here::here("./man/figures/svg/fig_6_95.jpg"), 
       dpi = 300,  plot=res_long_term_sites, width=25, height=14, units = "cm")
ggsave(filename = here::here("./man/figures/svg/fig_6_95.eps"), 
       dpi = 300,  plot=res_long_term_sites, width=25, height=14, units = "cm")
ggsave(filename = here::here("./man/figures/svg/fig_6_95.svg"), 
       dpi = 300,  plot=res_long_term_sites, width=25, height=14, units = "cm")

@@

sevRt_site95 <- plotrelsite(sdfsite95, v = "rt", x = 'd_severity', label = 'disturb_year', site = "site",
                 titulo = "Rt", coord_eqX = Inf, xLM = -Inf, yLM = 0.5) + xlab('') + ylab('')

sevRs_site95 <- plotrelsite(sdfsite95, v = "rs", x = 'd_severity', label = 'disturb_year', site = "site",
                 titulo = "Rs", coord_eqX = Inf, xLM = 10, yLM = 0.5) + xlab('') + ylab('')

sevRc_site95 <- plotrelsite(sdfsite95, v = "rc", x = 'd_severity', label = 'disturb_year', site = "site",
                 titulo = "Rc",  coord_eqX = -Inf, xLM = 10, yLM = 0.5) + xlab("Drought Severity") + ylab("")

@@

ggsave(filename = here::here("./man/figures/svg/fig_6_no95.jpg"), 
       dpi = 300,  plot=res_long_term_sites95, width=25, height=14, units = "cm")
ggsave(filename = here::here("./man/figures/svg/fig_6_no95.eps"), 
       dpi = 300,  plot=res_long_term_sites95, width=25, height=14, units = "cm")
ggsave(filename = here::here("./man/figures/svg/fig_6_no95.svg"), 
       dpi = 300,  plot=res_long_term_sites95, width=25, height=14, units = "cm")
```

# Check lm partial

```{r, eval= FALSE, echo = FALSE}
library(ggpubr)

ggscatter(
  sd, x = "Sepal.Length", y = "Sepal.Width",
  color = "Species", palette = "jco",
  add = "reg.line"
  ) +
  facet_wrap(~site) +
  stat_cor(label.y = 4.4) +
  stat_regline_equation(label.y = 4.2)

df1 <- sdfsite %>% dplyr::select(disturb_year, d_severity, rt_mean_mean, site)

  ggscatter(df1, x = "d_severity", y = "rt_mean_mean", 
            color = "site", palette = c("#00AFBB", "#E7B800", "#FC4E07"), 
            label = "disturb_year", font.label = c(8, "plain"), repel = TRUE, 
            add = "reg.line", ccor.coef = TRUE) +
    ggpubr::stat_cor(label.x= 15, aes(color = site, label = paste(..rr.label.., ..p.label.., sep = "~`,`~"))) +
    ggpubr::stat_cor(label.x= 15, aes(label = paste(..rr.label.., ..p.label.., sep = "~`,`~")))

df2 <-sdfsite %>% dplyr::select(disturb_year, site, rt_mean_mean, rs_mean_mean, rc_mean_mean, d_severity) %>%  gather(variable, value, c("rt_mean_mean", "rs_mean_mean", "rc_mean_mean"))

ggscatter(df2, x = "d_severity", y = "value", 
            color = "site", palette = c("#00AFBB", "#E7B800", "#FC4E07"), 
            label = "disturb_year", font.label = c(8, "plain"), repel = TRUE, 
            add = "reg.line", ccor.coef = TRUE) + 
  facet_wrap(~variable, scales = "free") + 
  ggpubr::stat_cor(label.x= 15, aes(color = site, label = paste(..rr.label.., ..p.label.., sep = "~`,`~"))) +
    ggpubr::stat_cor(label.x= 15, aes(label = paste(..rr.label.., ..p.label.., sep = "~`,`~")))
